---
title: "Conceptual Framework for Defections"
subtitle: "When and Why Do Regime Supporters Cease Their Support?"
author: "Jonathan Pinckney"
date: "2026-02-17"
format:
  revealjs:
    theme: [dark, custom.scss]
    slide-number: true
    width: 1280
    height: 720
    center: false
    hash-type: number
    footer: "Defections Project — Conceptual Framework"
    title-slide-attributes:
      data-background-color: "#1a1a2e"
---

```{r}
library(tidyverse)
```



## The Motivating Question {.smaller}

> When do supporters of authoritarian and democratic backsliding regimes cease to support their leaders? When do they [**defect**]{.highlight-red}?

. . .

- Defection is [**consequential**]{.highlight-red} — it shapes regime stability and transformation
- Yet we lack a [**unifying framework**]{.highlight-red} for understanding defection across contexts
- 

. . .

**Goal:** Develop a framework to guide research and inform practitioners seeking to advance, defend, and repair democracy.


## Power Rests on Obedience {.smaller}

All political regimes rely on key groups and institutions to provide power and resources [(Sharp 1973)]{.small-gray}:

. . .

- [**Security forces**]{.highlight-red} → capacity for repression
- [**Media**]{.highlight-red} → regime narratives and public compliance
- [**Citizens**]{.highlight-red} → taxes, obedience, legitimacy
- [**Business elites**]{.highlight-red} → economic resources
- [**Legislators & bureaucrats**]{.highlight-red} → institutional functioning

. . .




## What Is Defection? {.smaller}

[**Defection**]{.highlight-red} is any instance in which a regime supporter [**substantively decreases**]{.highlight-red} their level of regime support.

. . .

Examples:

- Soldiers refuse to open fire on demonstrators
- Government attorneys refuse politically-motivated prosecutions
- Legislators resign or vote against their party leadership
- Businesses withdraw financial support

. . .

Defection is especially crucial in processes of [**nonviolent resistance**]{.highlight-red}, whose logic rests not on eliminating opponents but on *shifting them to one's side* [(Chenoweth & Stephan 2011)]{.small-gray}.


## The Spectrum of Allies and Opponents {.smaller}

Adapted from [Oppenheimer & Lakey (1965)]{.small-gray}:

```{r}
#| fig-align: center
#| 
vlines <- tibble(x = 0:5)
spectrum.data <- tibble(
  y = rep(0,5),
  x = seq(0.5,4.5,1),
  label = c("Active\nOpposition","Passive\nOpposition","Neutral\n","Passive\nSupport","Active\nSupport")
)
arrows <- tibble(
  x = c(4.5,3.5,2.5),
  y = c(0.5,0.2,0.8),
  xend = c(2.5,1.5,0.5),
  yend = c(0.5,0.2,0.8),
  label = c("Defection","Defection","Not Defection")
)

# Color palette for dark background
col_opposition <- "#A93226"   # warm red
col_support    <- "#4ECDC4"   # teal
col_neutral    <- "#FFE66D"   # yellow
col_arrow_def  <- "#A93226"
col_arrow_ndef <- "#4ECDC4"
col_text       <- "white"
col_hline      <- "white"
col_vline      <- "white"

# Assign colors to spectrum labels
spectrum.data <- spectrum.data |>
  mutate(color = c(col_opposition, col_opposition, col_neutral, col_support, col_support))

# Assign colors to arrows
arrows <- arrows |>
  mutate(color = c(col_arrow_def, col_arrow_def, col_arrow_ndef))

ggplot(spectrum.data, aes(x, y, label = label)) +
  geom_text(
    aes(color = "white"),
    vjust = 1.5,
    fontface = "bold",
    size = 8
  ) +
  geom_hline(yintercept = 0, color = col_hline, linewidth = 0.8) +
  geom_vline(
    data = vlines,
    aes(xintercept = x),
    linetype = "dashed",
    color = col_vline,
    linewidth = 0.5
  ) +
  geom_segment(
    data = arrows,
    aes(x = x, y = y, xend = xend, yend = yend, color = color),
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    linewidth = 1.5
  ) +
  geom_text(
    data = arrows,
    aes(
      x = (x + xend) / 2,
      y = (y + yend) / 2,
      label = label,
      color = color
    ),
    vjust = -1,
    fontface = "bold",
    size = 8
  ) +
  scale_color_identity() +
  scale_y_continuous(limits = c(-0.4, 1.2)) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#191919", color = "#191919"),
    panel.background = element_rect(fill = "#191919", color = "#191919"),
    panel.border = element_blank(),
    text = element_text(size = 22)
  )
```

. . .

- [**Loyalty Shifts**]{.highlight-red} = any movement along the spectrum
- [**Defection**]{.highlight-red} = any leftward movement *beginning* in active or passive support


## Scope Conditions {.smaller}

This project focuses on defection within:

. . .

**1. Authoritarian and democratic backsliding regimes**

- Repression makes defection far riskier than in healthy democracies [(Davenport 2007)]{.small-gray}
- Different risk environments → different motivators
- Core motivation: understanding how to protect, repair, and advance democracy

. . .

**2. The contemporary world (post–Cold War)**

- Modern authoritarianism relies more on pseudo-democratic institutions, less on brute repression [(Guriev & Treisman 2019)]{.small-gray}
- Democratic backsliding today is gradual, centered on executive aggrandizement under a veneer of legality [(Bermeo 2016; Riedl et al. 2025)]{.small-gray}


## What Do We Already Know? {.smaller}

Different forms of defection have been studied in **isolation**:

:::: {.columns}

::: {.column width="50%"}

- **Military/security defection** during civil resistance [(Nepstad 2013; Grewal 2019; Dahl, Celestino & Gates 2026)]{.small-gray}
- **Legislative defection** from ruling parties in autocracies [(del Río 2022; Reuter & Gandhi 2011)]{.small-gray}

:::

::: {.column width="50%"}

- **Individual elite defection** from regime coalitions [(Hale & Colton 2017; Tertytchnaya 2020)]{.small-gray}
- **Mass voter defection** in electoral autocracies

:::

::::


## Three Consensus Findings {.smaller}

::: {style="display: grid; grid-template-rows: repeat(3, 1fr); height: 70vh;"}

::: {.fragment style="align-self: center; padding-left: 0%;"}
### Defection is [Consequential]{.highlight-red}
:::

::: {.fragment style="align-self: center; padding-left: 15%;"}
### Defection is [Difficult]{.highlight-red}
:::

::: {.fragment style="align-self: center; padding-left: 30%;"}
### Defection is [Complex]{.highlight-red} but [Predictable]{.highlight-red}
:::

:::

## {background-color="#1a1a2e"}

::: {style="text-align: center; margin-top: 150px;"}

[Guiding Assumptions]{style="font-size: 1.5em; font-weight: bold;"}

:::


## Defection is Individual or Collective {.smaller}

:::: {.columns}

::: {.column width="50%"}

Defection occurs at both levels:

::: {.incremental}
- [**Individual (micro):**]{.highlight-red} Personal decisions that may or may not affect one's group
- [**Group (meso):**]{.highlight-red} Collective defection that cannot be reduced to aggregated individual decisions
:::

[Research at only one level that doesn't accounr for the other yields an incomplete picture.]{.fragment}

:::

::: {.column width="50%"}

```{r}
#| fig-width: 5
#| fig-height: 6
#| fig-align: center
library(ggplot2)

# Define box positions
boxes <- data.frame(
  x = c(2.5, 2.5, 1, 2.5, 4),
  y = c(5, 3, 1, 1, 1),
  label = c("Political System", "Group", "Individual", "Individual", "Individual"),
  width = c(2.4, 1.8, 1.4, 1.4, 1.4),
  height = c(0.7, 0.7, 0.7, 0.7, 0.7),
  fill = rep("#A93226",5)
)

# Define connecting lines (from bottom of upper box to top of lower box)
lines <- data.frame(
  x =    c(2.5, 2.5, 2.5, 2.5),
  xend = c(2.5, 1.0, 2.5, 4.0),
  y =    c(5 - 0.35, 3 - 0.35, 3 - 0.35, 3 - 0.35),
  yend = c(3 + 0.35, 1 + 0.35, 1 + 0.35, 1 + 0.35)
)

ggplot() +
  # Draw connecting lines
  geom_segment(
    data = lines,
    aes(x = x, y = y, xend = xend, yend = yend),
    color = "#ffffff88",
    linewidth = 1.2
  ) +
  # Draw rounded rectangles
  geom_tile(
    data = boxes,
    aes(x = x, y = y, width = width, height = height, fill = fill),
    color = NA,
    linejoin = "round"
  ) +
  # Add labels
  geom_text(
    data = boxes,
    aes(x = x, y = y, label = label),
    color = "white",
    fontface = "bold",
    size = 5
  ) +
  scale_fill_identity() +
  coord_cartesian(xlim = c(-0.2, 5.2), ylim = c(0.2, 5.8)) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#191919", color = "#191919"),
    panel.background = element_rect(fill = "#191919", color = "#191919"),
  )
```

:::

::::


## Defection Is Social {.smaller}

:::: {.columns}

::: {.column width="50%"}

Understanding defection requires attention to:

- [**Embeddedness**]{.highlight-red} within broader social networks
- [**Information Flows**]{.highlight-red} through those networks

:::

::: {.column width="50%"}

```{r}
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
library(igraph)

set.seed(42)
edges <- data.frame(
  from = c(1, 1, 2, 2, 3, 3, 4, 5, 5, 6, 7, 8, 1, 4, 6),
  to   = c(2, 3, 4, 5, 6, 7, 5, 6, 8, 8, 8, 9, 6, 7, 9)
)
g <- graph_from_data_frame(edges, directed = TRUE)

node_colors <- c(
  "#4ECDC4", "#4ECDC4", "#4ECDC4", "#A93226", "#4ECDC4",
  "#A93226", "#A93226", "#4ECDC4", "#A93226"
)

coords <- layout_with_fr(g)

par(bg = NA, mar = c(0, 0, 0, 0))
plot(g,
  layout = coords,
  vertex.color = node_colors,
  vertex.frame.color = node_colors,
  vertex.size = 18,
  vertex.label = NA,
  edge.color = "#ffffff55",
  edge.arrow.size = 0.5,
  edge.width = 1.5
)
```

:::

::::


## Defection Has Multi-Faceted, Obscure Motivations {.smaller}

People defect for many reasons — often simultaneously:

- [**Rational self-interest**]{.highlight-red} (material utility maximization)
- [**Moral conviction**]{.highlight-red} [(Pearlman 2018)]{.small-gray}
- [**Changing social norms**]{.highlight-red} of appropriate behavior

. . .

These motivations interact:

- Material interests are easier to advance with moral justification
- Principled commitment is easier to sustain when materially beneficial

. . .

**Key caution:** Self-reported explanations are one data source among many, not definitive answers. People construct narratives to shape self and collective image [(Nisbett & Wilson 1977)]{.small-gray}.


## Defection Is Behavioral {.smaller}

Defection is defined not by what people [**think**]{.highlight-red} but by what they [**do**]{.highlight-red}.

. . .

- There must be an [**observable behavioral change**]{.highlight-red} vis-à-vis the regime
- Changes in beliefs or attitudes that don't produce behavior change ≠ defection

. . .

But speech counts as behavior!

- In authoritarian contexts, even slight indications of dissent are politically meaningful

. . .

**Caveat:** What counts as defection depends what constitutes [*loyalty*]{.highlight-red} for that individual or group.

- Havel's shopkeeper removing "Workers of the World, Unite" sign


## Defection Takes Many Forms {.smaller}

Two dimensions categorize forms of defection:

|  | [**Breaking**]{.highlight-red} (exit/outside) | [**Binding**]{.highlight-red} (voice/inside) |
|---|---|---|
| [**Speaking**]{.highlight-red} | Public condemnation | Private disapproval |
| [**Acting**]{.highlight-red} | Joining a demonstration | Creating internal reforms |
| [**Standing in the way**]{.highlight-red} | Disobeying orders | Foot-dragging |

. . .

- [**Breaking**]{.highlight-red} severs ties with the regime (visible, explicit)
- [**Binding**]{.highlight-red} maintains identity within regime while applying internal pressure (quiet, behind-the-scenes)
- Defection over time likely begins with **binding** → then **breaking**


## {background-color="#1a1a2e"}

::: {style="text-align: center; margin-top: 150px;"}

[Initial Propositions]{style="font-size: 1.5em; font-weight: bold;"}

:::


## Defection Requires a Way Out {.smaller}

Regime supporters won't defect without perceiving an **acceptable post-defection future**.

. . .

- Corrupt business elites fear both losing benefits *and* opposition animosity [(Sinanoglu 2025)]{.small-gray}
- Without viable post-defection pathways, defection is particularly rare [(Langston 2002)]{.small-gray}
- Eliminating alternatives is a key strategy of authoritarian control

. . .

**Implication for resistance campaigns:**

- Create alternative institutions
- Negotiate exit plans
- Promote an **inclusive vision** of a post-regime future where potential defectors can see themselves


## It's Not What You Know, It's Who You Know {.smaller}

Defection is driven by **social connections** and a "logic of appropriateness" [(March & Olsen 1998)]{.small-gray}.

. . .

- Positive connections to individuals/groups with lower regime loyalty → reduced loyalty, increased defection probability
- When group boundaries blur (supporters see their group as including opponents) → in-group loyalty no longer implies regime loyalty

. . .

**For resistance campaigns, defection is more likely when:**

1. The campaign increases [**social connections**]{.highlight-red} to regime supporters
2. The campaign signals that [**dissent is more widespread and durable**]{.highlight-red} than previously believed


## But Material Consequences Matter {.smaller}

Social connections are important but likely insufficient. Defection in authoritarian contexts is **high-risk, high-cost**, thus may only occur when material costs of defection are lower than costs of loyalty.

. . .

**Tension:** Social vs. material logics of sparking defection

*Example:* A business supporting an authoritarian leader

- [**Social**]{.highlight-red} approach: Patronize the business → build connections → social logic of appropriateness
- [**Material**]{.highlight-red} approach: Boycott → cut connections → heighten material costs of loyalty

. . .

Campaigns are likely to be effective when they [**sequence**]{.highlight-red} social and material strategies — material pressure may be most effective when seen as a temporary break in a relationship that can be restored.


## Step Size and Regime Type {.smaller}

[**Little steps are easier than big steps**]{.highlight-red}

- Most common defection: passive support → neutrality
- Jumps of multiple levels (active support → active opposition) are rare
- Reaching active opposition typically requires passing through each intermediate stage

. . .

[**But authoritarianism can make defection all-or-nothing**]{.highlight-red}

- Authoritarian regimes punish even modest loyalty declines with severe repression
- This accumulates large numbers of "false supporters" [(Kuran 1991)]{.small-gray}
- When they defect, they do so dramatically

. . .

**Hypothesis:** Defection scale is directly related to degree of authoritarianism. More open → small, frequent defections. More authoritarian → rare but large and abrupt.


## {background-color="#1a1a2e"}

::: {style="text-align: center; margin-top: 150px;"}

[Discussion]{style="font-size: 1.5em; font-weight: bold;"}

:::


## Is This Framework Helpful? {.smaller}

[**Do the concepts and definitions in this framework illuminate the relevant phenomena?**]{.highlight-red}

[**In particular: is it helpful to lump together small moves and big moves?**]{.highlight-red}

. . .
```{r}
#| fig-align: center
#| 
vlines <- tibble(x = 0:5)
spectrum.data <- tibble(
  y = rep(0,5),
  x = seq(0.5,4.5,1),
  label = c("Active\nOpposition","Passive\nOpposition","Neutral\n","Passive\nSupport","Active\nSupport")
)
arrows <- tibble(
  x = c(4.5,3.5,2.5),
  y = c(0.5,0.2,0.8),
  xend = c(3.5,1.5,0.5),
  yend = c(0.5,0.2,0.8),
  label = c("Defection","Defection","Not Defection")
)

# Color palette for dark background
col_opposition <- "#A93226"   # warm red
col_support    <- "#4ECDC4"   # teal
col_neutral    <- "#FFE66D"   # yellow
col_arrow_def  <- "#A93226"
col_arrow_ndef <- "#4ECDC4"
col_text       <- "white"
col_hline      <- "white"
col_vline      <- "white"

# Assign colors to spectrum labels
spectrum.data <- spectrum.data |>
  mutate(color = c(col_opposition, col_opposition, col_neutral, col_support, col_support))

# Assign colors to arrows
arrows <- arrows |>
  mutate(color = c(col_arrow_def, col_arrow_def, col_arrow_ndef))

ggplot(spectrum.data, aes(x, y, label = label)) +
  geom_text(
    aes(color = "white"),
    vjust = 1.5,
    fontface = "bold",
    size = 8
  ) +
  geom_hline(yintercept = 0, color = col_hline, linewidth = 0.8) +
  geom_vline(
    data = vlines,
    aes(xintercept = x),
    linetype = "dashed",
    color = col_vline,
    linewidth = 0.5
  ) +
  geom_segment(
    data = arrows,
    aes(x = x, y = y, xend = xend, yend = yend, color = color),
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    linewidth = 1.5
  ) +
  geom_text(
    data = arrows,
    aes(
      x = (x + xend) / 2,
      y = (y + yend) / 2,
      label = label,
      color = color
    ),
    vjust = -1,
    fontface = "bold",
    size = 8
  ) +
  scale_color_identity() +
  scale_y_continuous(limits = c(-0.4, 1.2)) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#191919", color = "#191919"),
    panel.background = element_rect(fill = "#191919", color = "#191919"),
    panel.border = element_blank(),
    text = element_text(size = 22)
  )
```


## Gains and Losses {.smaller}

[**What is gained and what is lost from thinking of defections in this high-level, unified way?**]{.highlight-red}

. . .

*Some considerations:*

- Gain: Cross-pollination of insights across fragmented literatures
- Gain: Unified language for practitioners and researchers
- Loss: Context-specific dynamics may be obscured
- Loss: Risk of over-generalizing across diverse phenomena


## Evaluating the Assumptions {.smaller}

[**Which assumptions are clear/unclear or helpful/unhelpful? Are there contradictions?**]{.highlight-red}

The six assumptions:

1. Defection can be [**individual or collective**]{.highlight-red}
2. Defection is [**social**]{.highlight-red}
3. Defection has [**multi-faceted, obscure motivations**]{.highlight-red}
4. Defection is [**behavioral**]{.highlight-red}
5. Defection [**takes many forms**]{.highlight-red}
6. Defection is [**contextual**]{.highlight-red}



## What's Missing? {.smaller}

[**What additional assumptions are necessary to understand defections?**]{.highlight-red}

. . .

*Some possible prompts:*

- Role of **emotions** (fear, hope, shame, anger) as distinct from rational calculation or moral conviction?
- **Temporality** — does timing of defection matter independently?
- **Reversibility** — can defectors return to support? How does this affect the framework?


## Other Critiques? {.smaller}


- What are the methodological challenges of studying defection given the framework?
- Are the scope conditions (authoritarian + backsliding regimes) too broad or too narrow?
- How does this framework relate to adjacent concepts (exit/voice/loyalty, methods of nonviolent action)?


## {background-color="#1a1a2e"}

::: {style="text-align: center; margin-top: 150px;"}

[Thank You!]{style="font-size: 1.5em; font-weight: bold;"}

:::