---
title: "Developing Defections Data"
subtitle: "Two Early-Stage Resources"
author: "Jonathan Pinckney, Julian Fortuna, and Suhan Kacholia"
date: "2026-02-20"
format:
  revealjs:
    theme: [dark, custom.scss]
    slide-number: true
    embed-resources: true
    width: 1280
    height: 720
    center: false
    hash-type: number
    footer: "Defections Project — Data Resources"
    title-slide-attributes:
      data-background-color: "#16213e"
bibliography: "../main-library.bib"
---

```{r}
library(tidyverse)
library(marginaleffects)

working.data <- read_rds("data/working-data.rds")

group.cyear.data <- working.data[[1]]
cyear.data <- working.data[[2]]
```

## Key Questions for New Data Sources


- How can these resources (and others) fit within a common [conceptual  framework]{.highlight-red}?
- How can we design these resources to most effectively [answer questions]{.highlight-red} from the previous session?
- What methodological [challenges]{.highlight-red} should we be keeping in mind?
- How can we design these resources to most effectively [pipeline findings]{.highlight-red} to relevant practitioner audiences?

## What Data Resources Already Exist?

- Military and civil service defections during civil resistance campaigns [@dahl-disaggregating-2026;@chenoweth-updating-2022]
- Loyalty/defection at the pillar of support level for some pillars in some backsliding spells [@pinckney-breaking-2025]
- Individual-level data on authoritarian legislators/ministers including instances of defection to the opposition [@delrio-strategic-2022] 

## What is Missing?

::: {.incremental}
- Data that tracks [broader populations]{.highlight-red} of regime supporters across time and contexts
- Data on a wider range of types of defections, recognizing the importance of [context]{.highlight-red} and [subtle/quiet action]{.highlight-red}
- Comparative case studies and stories that [process-trace]{.highlight-red} defections across cases and contexts
:::

## {background-color="#16213e"}

::: {style="text-align: center; margin-top: 150px;"}

[Cross-National Defections Data]{style="font-size: 1.5em; font-weight: bold;"}

:::

## What is This Resource For?

:::{.incremental style="line-height: 2.5; margin-top: 30px;"}
- [Large-*n* analysis]{.highlight-red} to identify factors that predict defection cross-nationally
- Identifying key global cases of defection for [process-tracing]{.highlight-red}
:::

## Scope Conditions

Every authoritarian and democratic backsliding regime in the post-Cold War era - `r nrow(filter(cyear.data,v2x_regime < 2 | in_manifest_episode == TRUE))` country-years [@luhrmann-regimes-2018; @luhrmann-third-2019]

.  .  .

```{r}
#| fig-align: "center"
#| out-width: "90%"
#| fig-width: 11.5
#| fig-height: 5
#| fig-cap: "Regime Periods in Hungary 1990-2024"

hungary.data <- cyear.data %>%
  select(country_name,year,v2x_polyarchy) %>% 
  filter(country_name == "Hungary" & year > 1989)

regimes <- tibble(xmin = c(1990,2010,2017),
                  xmax = c(2010,2017,2024),
                  Regime = c("Democracy","Backsliding","Authoritarian"))

ggplot() +
  geom_rect(data = regimes,
            aes(xmin =xmin,xmax = xmax,ymin = -Inf,ymax = Inf,fill = Regime),
            alpha = 0.4) +
  geom_line(data = hungary.data,
            aes(x = year,y = v2x_polyarchy),
            color = "white", linewidth = 1) +
  scale_fill_manual(name = "",
                    values = c("Democracy" = "#2196F3",
                               "Backsliding" = "#FFC107",
                               "Authoritarian" = "#F44336"),
                    guide = guide_legend(reverse = T)) +
  labs(x = "Year",y = "Electoral Democracy Score") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey40"),
    plot.background = element_rect(fill = "#191919", color = "#191919"),
    panel.background = element_rect(fill = "#191919", color = "#191919"),
    panel.border = element_blank(),
        text = element_text(size = 24,color = "white"),
        axis.text = element_text(color = "grey80"),
        legend.text = element_text(color = "white"))
```

## Unit of Analysis

The [**regime support group**]{.highlight-red}-country-year

> "A group of individuals who are supportive of the existing regime, and, *if* it were to retract support would substantially increase the chance that the regime would lose power" [@knutsen-throne-2025, 5]

. .  .  

The *Regime Support and Opposition Groups* dataset: country expert data on whether 14 groups [*support*, *oppose*, or *actively oppose*]{.highlight-red} a political regime.

.  .  .  

Variables are measured as percentage of experts who agree that the group falls into that category

## 

:::: {.columns}

::: {.column width="30%"}



|[Groups]{.highlight-red}|
|-----------|
::: {style="font-size: 0.651em"}
Aristocracy  
Landholders  
Party Elites  
Business Elites  
Bureaucracy  
[Military]{.highlight-red}  
Ethnic Group  
Religious Group  
Local Elites  
Urban Workers  
Urban Middle Class  
Rural Workers  
Rural Middle Class  
Foreign Government  
:::


:::

::: {.column width="70%"}

```{r}
#| fig-width: 7
#| fig-height: 5.5

thailand.mil <- group.cyear.data %>%
  filter(country_name == "Burma/Myanmar" & group == "Military" & year > 1989) %>%
  select(year, support, oppose, active.oppose) %>%
  pivot_longer(cols = c(support, oppose, active.oppose),
               names_to = "measure", values_to = "value") %>%
  mutate(measure = factor(measure,
                          levels = c("support", "oppose", "active.oppose"),
                          labels = c("Support", "Oppose", "Actively Oppose")))

ggplot(thailand.mil, aes(x = year, y = value, color = measure)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(name = "",
                     values = c("Support" = "#2196F3",
                                "Oppose" = "#FFC107",
                                "Actively Oppose" = "#F44336")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1)) +
  labs(x = "Year", y = "Expert Consensus",
       title = "Military — Burma/Myanmar") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey40"),
        plot.background = element_rect(fill = "#191919", color = "#191919"),
        panel.background = element_rect(fill = "#191919", color = "#191919"),
        panel.border = element_blank(),
        text = element_text(size = 18, color = "white"),
        axis.text = element_text(color = "grey80"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white", size = 20))
```


:::

::::

## Potential Defection Years

Country-years in which consensus on regime support was above a threshold in the prior year, and consensus on [*support falls*]{.highlight-red} below a threshold or [*opposition rises*]{.highlight-red} above a threshold in the current year.

```{r}
#| fig-align: "center"
#| out-width: "90%"
#| fig-width: 11.5
#| fig-height: 5.5

potential.dfx.plot <- group.cyear.data %>%
  filter(year > 1989 & country_name == "Burma/Myanmar") %>%
  # group_by(group) %>%
  # filter(max.support > 0.5 | group %in% c("Military","Bureaucracy")) %>%
  # ungroup() %>%
  mutate(potential.dfx = factor(potential.dfx, levels = c(0, 1),
                                labels = c("No", "Yes")))

ggplot(potential.dfx.plot, aes(x = year, y = group, fill = potential.dfx)) +
  geom_tile(color = "#191919", linewidth = 0.5) +
  scale_fill_manual(name = "Potential Defection",
                    values = c("No" = "#1565C0", "Yes" = "#C41C3B"),
                    na.value = "grey30") +
  # scale_x_continuous(breaks = seq(1992, 2024, by = 4)) +
  labs(x = "Year", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top",
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#191919", color = "#191919"),
        panel.background = element_rect(fill = "#191919", color = "#191919"),
        panel.border = element_blank(),
        text = element_text(size = 24, color = "white"),
        axis.text = element_text(color = "grey80"),
        axis.text.y = element_text(size = 16),
        legend.text = element_text(color = "white"))
```



## Potential Variables {.smaller}

Limit to variables not readily available from existing data sources

```{=html}
<table style="width: 90%; margin: auto; border-collapse: separate; border-spacing: 0; font-size: 0.75em;">
  <thead>
    <tr style="border-bottom: 2px solid #888;">
      <th style="text-align: left; padding: 12px 16px; color: #FFC107; width: 30%;">Variable</th>
      <th style="text-align: left; padding: 12px 16px; color: #ccc;">Question</th>
    </tr>
  </thead>
  <tbody>
    <tr style="border-bottom: 1px solid #444;">
      <td style="padding: 10px 16px; color: #F44336; font-family: monospace; font-weight: bold;">group.loyalty.unity</td>
      <td style="padding: 10px 16px; color: #eee;">How united is this group in its allegiance or opposition to the regime?</td>
    </tr>
    <tr style="border-bottom: 1px solid #444;">
      <td style="padding: 10px 16px; color: #F44336; font-family: monospace; font-weight: bold;">group.loyalty.intensity</td>
      <td style="padding: 10px 16px; color: #eee;">What is the highest level of opposition to the regime that you observe among this group?</td>
    </tr>
    <tr style="border-bottom: 1px solid #444;">
      <td style="padding: 10px 16px; color: #F44336; font-family: monospace; font-weight: bold;">opposition.tactics</td>
      <td style="padding: 10px 16px; color: #eee;">Did the opposition engage in any intentional attempts to spark defections from this group?</td>
    </tr>
    <tr style="border-bottom: 1px solid #444;">
      <td style="padding: 10px 16px; color: #F44336; font-family: monospace; font-weight: bold;">opposition.tactics.list</td>
      <td style="padding: 10px 16px; color: #eee;">What specific tactics did the opposition employ to attempt to spark defections from this group?</td>
    </tr>
    <tr>
      <td style="padding: 10px 16px; color: #F44336; font-family: monospace; font-weight: bold;">defection.tactics</td>
      <td style="padding: 10px 16px; color: #eee;">What specific actions did members of this support group engage in that resisted or opposed the regime in power?</td>
    </tr>
     <tr>
      <td style="padding: 10px 16px; color: #F44336; font-family: monospace; font-weight: bold;">social.connection</td>
      <td style="padding: 10px 16px; color: #eee;">How dense are the social connections between this group and the opposition?</td>
    </tr>
     <tr>
      <td style="padding: 10px 16px; color: #F44336; font-family: monospace; font-weight: bold;">vignette</td>
      <td style="padding: 10px 16px; color: #eee;">Is this group in this year a good candidate for a defection case study?</td>
    </tr>

  </tbody>
</table>
```

## Data Collection and Coding Process

::: {.incremental}
- Code variables for each [*relevant*]{.highlight-red} support group beginning from first year of a regime
- Carry values forward until [*potential defection year*]{.highlight-red}, then carry values forward until next potential defection year
- Code based on curated list of scholarly and media sources. Double-code with human RA coder and LLM-based AI agent
- Where possible, check against existing datasets to see what is being captured and what is not
:::

## Some Preliminary Modeling!

. . .

[Civil resistance campaigns are associated with more potential defections!]{.highlight-red}

```{r}
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 5

# log.mod <- glm(support.drop ~ NONVIOL + v2x_polyarchy + as_factor(year) + country_name, family = binomial(link = "logit"),
#                data = cyear.data)
# 
# p <- plot_predictions(log.mod, by = "NONVIOL")
# 
# write_rds(p,"predplot.rds")
p <- read_rds("predplot.rds")

p$layers <- lapply(p$layers, function(layer) {
  layer$aes_params$colour <- "#A93226"
  layer$aes_params$fill   <- "#A93226"
  layer$aes_params$size <- 2
  layer$aes_params$linewidth = 2
  layer
})

p +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(labels = c("No", "Yes")) +
  labs(x = "Civil Resistance Campaign", y = "Defection Probability") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major = element_line(color = "grey40"),
    plot.background  = element_rect(fill = "#191919", color = "#191919"),
    panel.background = element_rect(fill = "#191919", color = "#191919"),
    panel.border     = element_blank(),
    text             = element_text(size = 20, color = "white"),
    axis.text        = element_text(color = "grey80", size = 18)
  )

```



## {background-color="#16213e"}

::: {style="text-align: center; margin-top: 150px;"}

### Discussion

:::

## A Few Major Questions for Feedback

- What are the biggest [problems]{.highlight-red} with this approach? What is it likely to miss?
- What [variables]{.highlight-red} should we include in order to test hypotheses/provide insights on the questions brought up in the second session?
- What should we be looking for in informing the [*vignette*]{.highlight-red} variable? What kinds of defection stories from a global context should we be flagging?







## {background-color="#16213e"}

::: {style="text-align: center; margin-top: 150px;"}

### Now over to Julian and Suhan!

:::

## References